---
title: "The freqlist function"
author: "Tina Gunderson"
date: '`r format(Sys.time(),"%d %B, %Y")`'
output:
  rmarkdown::html_vignette:
    toc: yes
    toc_depth: 3
header-includes: \usepackage{tabularx}
vignette: |
  %\VignetteIndexEntry{The freqlist function}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, tidy.opts=list(width.cutoff=80), tidy=TRUE,  comment=NA)
options(width=80, max.print=1000)
```

\newpage

# Overview 

`freqlist` is a function meant to produce output similar to SAS's `PROC FREQ` procedure when using the `/list` option of the `TABLE` statement.
`freqlist` provides options for handling missing or sparse data and can provide cumulative counts and percentages based on subgroups.
It depends on the `knitr` package for printing. 

```{r message = FALSE}
require(arsenal)
```

## Sample dataset

For our examples, we'll load the `mockstudy` data included with this package and use it to create a basic table.
Because they have fewer levels, for brevity, we'll use the variables arm, sex, and mdquality.s to create the example table.
We'll retain NAs in the table creation.
See the appendix for notes regarding default NA handling and other useful information regarding tables in R. 

```{r loading.data}
# load the data
data(mockstudy)

# examine the data
str(mockstudy)

# retain NAs when creating the table using the useNA argument
tab.ex <- table(mockstudy[, c("arm", "sex", "mdquality.s")], useNA="ifany")
```

\newpage
 
# The `freqlist` object

The `freqlist` function returns an object of class `freqlist`, which has three parts: `freqlist`, `byVar`, and `labels`.

 - `freqlist` is a single data frame containing all contingency tables with calculated frequencies, cumulative frequencies, percentages, and cumulative percentages.
 
 - `byVar` and `labels` are used in the `summary` method for subgroups and variable names, which will be covered in later examples.

```{r console.output}
noby <- freqlist(tab.ex)

str(noby)

# view the data frame portion of freqlist output
noby[["freqlist"]] ## or use as.data.frame(noby)
```

\newpage

# Basic output using `summary` 

The `summary` method for `freqlist` relies on the `kable` function (in the `knitr` package) for printing.
`knitr::kable` converts the output to markdown which can be printed in the console or easily rendered in Word, pdf, or html documents.

Note that you must supply `results="asis"` to properly format the markdown output.

```{r, results = 'asis'}
summary(noby)
```

Additional arguments (except digits) in the `kable` function can be passed through. Perhaps the most useful is `caption`.

```{r, results = 'asis'}
summary(noby, caption="Basic freqlist output")
```

You can also easily pull out the `freqlist` data frame for more complicated formatting or manipulation
(e.g. with another function such as `xtable` or `pander`) using `as.data.frame`:

```{r}
head(as.data.frame(noby))
```

\newpage

# Rounding percentage digits or changing variable names for printing

The digits argument takes a single numeric value and controls the rounding of percentages in the output.
The labelTranslations argument is a character vector whose length must be equal to the number of factors used in the table.
Note: this does not change the names of the data frame in the freqlist object, only those used in printing.
Both options are applied in the following example. 

```{r labelTranslations, results = 'asis'}

withnames <- freqlist(tab.ex, labelTranslations = c("Treatment Arm","Gender","LASA QOL"), digits = 0)

summary(withnames)

```

\newpage

# Additional examples

## Including combinations with frequencies of zero

The sparse argument takes a single logical value as input. The default option is FALSE. If set to TRUE, the sparse option will include combinations with frequencies of zero in the list of results. As our initial table did not have any such levels, we create a second table to use in our example.

```{r sparse, results = 'asis'}
# we create a second table example to showcase the sparse argument
tab.sparse <- table(mockstudy[, c("race","sex","arm")])

nobysparse <- freqlist(tab.sparse, sparse = TRUE, digits=1)
summary(nobysparse)
```

## Options for NA handling

The various na.options allow you to include or exclude data with missing values for one or more factor levels in the counts and percentages as well as show the missing data but exclude it from the cumulative counts and percentages. The default option is to include all combinations with missing values. 

```{r na.options, results = 'asis'}
summary(freqlist(tab.ex, na.options="include"))
summary(freqlist(tab.ex, na.options="showexclude"))
summary(freqlist(tab.ex, na.options="remove"))
```

## Frequency counts and percentages subset by factor levels 

The groupBy argument internally subsets the data by the specified factor prior to calculating cumulative counts and percentages.
By default, when used each subset will print in a separate table. Using the `single = TRUE` option when printing will collapse
the subsetted result into a single table.

```{r freq.counts, results='asis'}
withby <- freqlist(tab.ex, groupBy = c("arm","sex"))
summary(withby)

#using the single = TRUE argument will collapse results into a single table for printing
summary(withby, single = TRUE)

```

## Change labels on the fly

At this time, the labels can be changed just for the variables (e.g. not the frequency columns).

```{r changelabs, results = 'asis'}
labels(noby) <- c("Arm", "Sex", "OtherThing")
summary(noby)
```

You can also supply `labelTranslations` to `summary`.

```{r, results = 'asis'}
summary(noby, labelTranslations = c("Hi there", "What up", "Bye"))
```

## Using `xtable` to format and print `freqlist` results

Fair warning: `xtable` has kind of a steep learning curve. These examples are given without explanation for more advanced users.

```{r xtable.setup}
require(xtable)

# set up custom function for xtable text
italic <- function(x){
paste0('<i>', x, '</i>')
}

```

```{r printxtable, results='asis'}
xftbl <- xtable(noby[["freqlist"]], 
  caption = "xtable formatted output of freqlist data frame", align="|r|r|r|r|c|c|c|r|")

# change the column names
names(xftbl)[1:3] <- c("Arm", "Gender", "LASA QOL")

print(xftbl, sanitize.colnames.function = italic, include.rownames = FALSE, type = "html", comment = FALSE)
```

\newpage

# Appendix: Notes regarding table options in R

## NAs

There are several widely used options for basic tables in R. The `table` function in base R is probably the most common;
by default it excludes NA values. You can change NA handling in `base::table` using the useNA or exclude arguments.

```{r}
# base table default removes NAs
tab.d1 <- base::table(mockstudy[, c("arm", "sex", "mdquality.s")], useNA="ifany")
tab.d1
```

`xtabs` is similar to `table`, but uses a formula-based syntax. However, there is not an option for retaining NAs in the
`xtabs` function; instead, NAs must be added to each level of the factor where present using the `addNA` function.

```{r}
# without specifying addNA
tab.d2 <- xtabs(formula = ~ arm + sex + mdquality.s, data = mockstudy)
tab.d2

# now with addNA
tab.d3 <- xtabs(~ arm + sex + addNA(mdquality.s), data = mockstudy)
tab.d3

```

## Table dimname names (dnn)

Supplying a data.frame to the `table` function without giving columns individually will create a contingency table
using all variables in the data.frame.

However, if the columns of a data.frame or matrix are supplied separately (i.e., as vectors),
column names will not be preserved.

```{r}
# providing variables separately (as vectors) drops column names
tab.d4 <- base::table(mockstudy[, "arm"], mockstudy[, "sex"], mockstudy[, "mdquality.s"])
tab.d4
```

If desired, you can use the `dnn` argument to pass variable names.

```{r}
# add the column name labels back using dnn option in base::table
tab.dnn <- base::table(mockstudy[, "arm"], mockstudy[, "sex"], mockstudy[, "mdquality.s"], dnn=c("Amy", "Susan", "George"))
tab.dnn
```

If using `freqlist`, you can provide the labels directly to `freqlist` or to `summary` using `labelTranslations`.
