---
title: "A Few Notes on Labels"
author: "Ethan Heinzen"
date: '`r format(Sys.time(),"%d %B, %Y")`'
output:
  rmarkdown::html_vignette:
    toc: true
vignette: |
  %\VignetteIndexEntry{A Few Notes on Labels}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
---

```{r include = FALSE}
knitr::opts_chunk$set(message = FALSE, results = 'asis')
```

# Introduction

The `arsenal` package relies somewhat heavily on variable labels to make output more "pretty".
A `label` here is understood to be a single character string with "pretty" text (i.e., not an "ugly" variable name).
Three of the main `arsenal` function use labels in their `summary()` output. There are several ways to set these labels.

We'll use the `mockstudy` dataset for all examples here:

```{r}
library(arsenal)
data(mockstudy)
library(magrittr)

# for 'freqlist' examples
tab.ex <- table(mockstudy[c("arm", "sex", "mdquality.s")], useNA="ifany")
```

# Examples

## Set labels in the function call

The `summary()` method for `tableby()`, `modelsum()`, and `freqlist()` objects contains a `labelTranslations = ` argument to specify labels
in the function call. Note that the `freqlist()` function matches labels in order, whereas the other two match labels by name. The labels
can be input as a list or a character vector.

```{r}
summary(freqlist(tab.ex),
        labelTranslations = c(arm = "Treatment Arm", sex = "Gender", mdquality.s = "LASA QOL"))
summary(tableby(arm ~ sex + age, data = mockstudy),
        labelTranslations = c(sex = "SEX", age = "Age, yrs"))
summary(modelsum(bmi ~ age, adjust = ~sex, data = mockstudy),
        labelTranslations = list(sexFemale = "Female", age = "Age, yrs"))
```

## Modify labels after the fact

Another option is to add labels after you have created the object. To do this, you can use the form
`labels(x) <- value` or use the pipe-able version, `set_labels()`.

```{r}
# the non-pipe version; somewhat clunky
tmp <- freqlist(tab.ex)
labels(tmp) <- c(arm = "Treatment Arm", sex = "Gender", mdquality.s = "LASA QOL")
summary(tmp)

# piped--much cleaner
mockstudy %>% 
  tableby(arm ~ sex + age, data = .) %>% 
  set_labels(c(sex = "SEX", age = "Age, yrs")) %>% 
  summary()

mockstudy %>% 
  modelsum(bmi ~ age, adjust = ~ sex, data = .) %>% 
  set_labels(list(sexFemale = "Female", age = "Age, yrs")) %>% 
  summary()
```

## Add labels to a `data.frame`

`tableby()` and `modelsum()` also allow you to have label attributes on the data. Note
that by default these attributes usually get dropped upon subsetting, but `tableby()` and
`modelsum()` use the `keep.labels()` function to retain them.

```{r}
mockstudy.lab <- keep.labels(mockstudy)
class(mockstudy$age)
class(mockstudy.lab$age)
```

To undo this, simply `loosen.labels()`:

```{r}
class(loosen.labels(mockstudy.lab)$age)
```

You can set attributes one at a time in two ways:

```{r}
attr(mockstudy.lab$sex, "label") <- "Sex"
labels(mockstudy.lab$age) <- "Age, yrs"
```

...or all at once:

```{r}
labels(mockstudy.lab) <- list(sex = "Sex", age = "Age, yrs")
summary(tableby(arm ~ sex + age, data = mockstudy.lab))
```

You can pipe this, too.

```{r}
mockstudy %>% 
  set_labels(list(sex = "SEX", age = "Age, yrs")) %>% 
  modelsum(bmi ~ age, adjust = ~ sex, data = .) %>% 
  summary()
```

To extract labels from a `data.frame`, simply use the `labels()` function:

```{r results='markdown'}
labels(mockstudy.lab)
```


