---
title: "The paired function"
author: "Ethan Heinzen, Beth Atkinson, Jason Sinnwell"
date: '`r format(Sys.time(),"%d %B, %Y")`'
output: 
  rmarkdown::html_vignette:
    toc: yes
    toc_depth: 3
vignette: |
  %\VignetteIndexEntry{The paired function}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
---

```{r echo = FALSE}
options(width = 100)
```

# Introduction

Another one of the most common tables in medical literature includes summary statistics for a set of variables paired across
two time points.  Locally at Mayo, the SAS macro `%paired` was written to create summary tables with a single call.
With the increasing interest in R, we have developed the function `paired()` to create similar tables within the R environment.  

This vignette is light on purpose; `paired()` piggybacks off of tableby, so most documentation there applies here, too.

# Simple Example

The first step when using the `paired()` function is to load the `arsenal` package.  We can't use `mockstudy` here because
we need a dataset with paired observations, so we'll create our own dataset.

```{r, load-data}
library(arsenal)
dat <- data.frame(
  tp = paste0("Time Point ", c(1, 2, 1, 2, 1, 2, 1, 2, 1, 2)),
  id = c(1, 1, 2, 2, 3, 3, 4, 4, 5, 6),
  Cat = c("A", "A", "A", "B", "B", "B", "B", "A", NA, "B"),
  Fac = factor(c("A", "B", "C", "A", "B", "C", "A", "B", "C", "A")),
  Num = c(1, 2, 3, 4, 4, 3, 3, 4, 0, NA),
  Ord = ordered(c("I", "II", "II", "III", "III", "III", "I", "III", "II", "I")),
  Lgl = c(TRUE, TRUE, FALSE, TRUE, FALSE, TRUE, TRUE, FALSE, FALSE, FALSE),
  Dat = as.Date("2018-05-01") + c(1, 1, 2, 2, 3, 4, 5, 6, 3, 4),
  stringsAsFactors = FALSE
)
```

To create a simple table stratified by time point, use a `formula=` statement to specify the variables that you want summarized and
the `id=` argument to specify the paired observations.

```{r results = 'asis'}
p <- paired(tp ~ Cat + Fac + Num + Ord + Lgl + Dat, data = dat, id = id, signed.rank.exact = FALSE)
summary(p)
```

The third column shows the difference between time point 1 and time point 2. For categorical variables, it reports
the percent of observations from time point 1 which changed in time point 2.

# NAs

Note that by default, observations which do not have both timepoints are removed.
This is easily changed using the `na.action = na.paired("<arg>")` argument. For example:

```{r results = 'asis'}
p <- paired(tp ~ Cat + Fac + Num + Ord + Lgl + Dat, data = dat, id = id,
            signed.rank.exact = FALSE, na.action = na.paired("fill"))
summary(p)
```

For more details, see the help page for `na.paired()`.

# Available Function Options

## Testing options

The tests used to calculate p-values differ by the variable type, but can be specified
explicitly in the formula statement or in the control function.

The following tests are accepted:

* `paired.t`: A paired t-test. 

* `mcnemar`: McNemar's test.

* `signed.rank`: the signed-rank test.

* `sign.test`: the sign test.

* `notest`: Don't perform a test.

## `paired.control` settings

A quick way to see what arguments are possible to utilize in a function is to use the `args()` command. Settings involving the number of digits can be set in `paired.control` or in `summary.tableby`.

```{r}
args(paired.control)
```

## `summary.tableby` settings

Since the "paired" object inherits "tableby", the `summary.tableby` function is what's actually used
to format and print the table.

```{r}
args(arsenal:::summary.tableby)
```

