---
title: "The write2 function"
author: "Ethan Heinzen"
date: '`r format(Sys.time(),"%d %B, %Y")`'
output:
  rmarkdown::html_vignette:
    toc: yes
    toc_depth: 3
vignette: |
  %\VignetteIndexEntry{The write2 function}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
---

```{r include = FALSE}
knitr::opts_chunk$set(eval = FALSE)
```

# Introduction

The `write2*()` functions were designed as an alternative to SAS's `ODS` procedure for useRs who want to save R Markdown tables
to separate Word, HTML, or PDF files without needing separate R Markdown programs.

There are three shortcut functions for the most common output types: HTML, PDF, and Word. Each of these three functions calls
`write2()`, an S3 function which accepts many file output types (see the help pages for `rmarkdown::render()`). Methods have been implemented for
`tableby()`, `modelsum()`, and `freqlist()`, but also `knitr::kable()`, `xtable::xtable()`, and `pander::pander_return()`.

The two most important things to recognize with `write2()` are the following:

1. Which function is being used to output the object. Sometimes the `write2` functions use `summary()`, while other times they will use `print()`.
   The details for each object specifically are described below.
   
2. How the `...` arguments are passed. To change the options for the summary-like or print-like function, you can pass named arguments which will
   in turn get passed to the appropriate function. Details for each object specifically are described below.

# Examples Using `arsenal` Objects

```{r}
library(arsenal)
data(mockstudy)
tmpdir <- tempdir()
```

## `tableby`

For `tableby` objects, the output function in `write2()` is `summary()`. For available arguments, see the help pages for `summary.tableby()`.
Don't use the option `text = TRUE` with the `write2` functions.

```{r}
mylabels <- list(sex = "SEX", age ="Age, yrs")
tab1 <- tableby(arm ~ sex + age, data=mockstudy)

write2html(tab1, paste0(tmpdir, "/test.tableby.html"), quiet = TRUE,
           title = "My test table",      # passed to summary.tableby
           labelTranslations = mylabels, # passed to summary.tableby
           total = FALSE                 # passed to summary.tableby
           )
```

## `modelsum`

For `modelsum` objects, the output function in `write2()` is `summary()`. For available arguments, see the help pages for `summary.modelsum()`.
Don't use the option `text = TRUE` with the `write2` functions.

```{r}
tab2 <- modelsum(alk.phos ~ arm + ps + hgb, adjust= ~ age + sex, family = "gaussian", data = mockstudy)

write2pdf(tab2, paste0(tmpdir, "/test.modelsum.pdf"), quiet = TRUE,
          title = "My test table", # passed to summary.modelsum
          show.intercept = FALSE,  # passed to summary.modelsum
          digits = 5               # passed to summary.modelsum
          )
```

## `freqlist`

For `freqlist` objects, the output function in `write2()` is `summary()`. For available arguments, see the help pages for `summary.freqlist()`.

```{r}
tab3 <- freqlist(table(mockstudy[, c("arm", "sex", "mdquality.s")], useNA = "ifany"), groupBy = c("arm", "sex"))

write2word(tab3, paste0(tmpdir, "/test.freqlist.doc"), quiet = TRUE,
           single = FALSE # passed to summary.freqlist
           )
```

# Examples Using Other Objects

## `knitr::kable()`

For objects resulting from a call to `kable()`, the output function in `write2()` is `print()`. There aren't any arguments to the `print.knitr_kable()` function.

```{r}
write2html(knitr::kable(head(mockstudy)), paste0(tmpdir, "/test.kable.html"), quiet = TRUE)
```

## `xtable::xtable()`

For `xtable` objects, the output function in `write2()` is `print()`. For available arguments, see the help pages for `print.xtable()`.

```{r}
write2pdf(xtable::xtable(head(mockstudy), caption = "My xtable"), paste0(tmpdir, "/test.xtable.pdf"), quiet = TRUE,
          comment = FALSE, # passed to print.xtable to turn off the default message about xtable version
          include.rownames = FALSE, # passed to print.xtable
          caption.placement = "top" # passed to print.xtable
          )
```

To make an HTML document, use the `print.xtable()` option `type = "html"`.

```{r}
write2html(xtable::xtable(head(mockstudy), caption = "My xtable"), paste0(tmpdir, "/test.xtable.html"), quiet = TRUE,
           type = "html",            # passed to print.xtable
           comment = FALSE, # passed to print.xtable to turn off the default message about xtable version
           include.rownames = FALSE, # passed to print.xtable
           caption.placement = "top" # passed to print.xtable
           )
```

User beware! `xtable()` is not compatible with `write2word()`.

## `pander::pander_return()`

Pander is a little bit more tricky. Since `pander::pander()` doesn't return an object, the useR should instead use
`pander::pander_return()`. For this (and for all character vectors), the the output function in `write2()` is `cat(sep = '\n')`.

```{r}
write2word(pander::pander_return(head(mockstudy)), file = paste0(tmpdir, "/test.pander.doc"), quiet = TRUE)
```

# FAQs

## How do I suppress the note about my document getting rendered?

This is easily accomplished by using the argument `quiet = TRUE` (passed to the `rmarkdown::render()` function).

```{r}
write2html(knitr::kable(head(mockstudy)), paste0(tmpdir, "/test.kable.quiet.html"),
           quiet = TRUE # passed to rmarkdown::render
           )
```

## How do I look at the temporary `.md` file?

This is easily accomplished by using the option `keep.md = TRUE`.

```{r}
write2html(knitr::kable(head(mockstudy)), paste0(tmpdir, "/test.kable.keep.md.html"),
           quiet = TRUE, # passed to rmarkdown::render
           keep.md = TRUE
           )
```

## How do I tweak the default format from `write2word()`, `write2html()`, or `write2pdf()`?

You can pass arguments to the format functions used behind the scenes.

```{r}
write2html(knitr::kable(head(mockstudy)), paste0(tmpdir, "/test.kable.theme.html"),
           quiet = TRUE,  # passed to rmarkdown::render
           theme = "yeti" # passed to rmarkdown::html_document
           )
```

See the help pages for `rmarkdown::word_document()`, `rmarkdown::html_document()`, and `rmarkdown::pdf_document()`.

## How do I output to a file format other than word, HTML, and PDF?

This can be done using the generic `write2()` function. The last argument in the function can be another format specification.
For details on the acceptable inputs, see the help page for `write2()`.

```{r}
write2(knitr::kable(head(mockstudy[, 1:4])), paste0(tmpdir, "/test.kable.rtf"),
       quiet = TRUE,  # passed to rmarkdown::render
       output_format = rmarkdown::rtf_document
       )
```
